LEGWORK: A python package for computing the evolution and detectability \\ of stellar-origin gravitational-wave sources with space-based detectors

### 2022-01-29 src/ms.tex
add PN clarification

The planned space-based gravitational wave detector LISA (Laser Interferometer Space Antenna) will present an entirely new view of gravitational waves by focusing on lower frequencies ($10^{-5} < f / \unit{Hz} < 10^{-1}$) than ground-based detectors. This will enable the study of many new source classes including mergers of supermassive black holes \citep[e.g.][]{Begelman+1980, Klein+2016, Bellovary2019}, extreme mass ratio inspirals \citep[e.g.,][]{Berti2006, Barack2007, Babak2017, Moore2017}, and cosmological GW backgrounds \citep[e.g.,][]{Bartolo2016, Caprini2016, Caldwell2019}. However, this frequency regime is also of interest for the detection of local stellar-mass binaries during their inspiral phase. LISA is expected to detect Galactic stellar-origin binaries containing combinations of white dwarfs, neutron stars, and black holes, ranging from the numerous double white dwarf population, to the rare but loud double black hole population.

@@

\lw{} is an open-source Python package designed to streamline the process of making predictions of LISA detection rates for stellar-origin binaries such that it is as fast, reliable and simple as possible. With \lw{} one can evolve the orbits of a binary or a collection of binaries and calculate their strain amplitudes for any range of frequency harmonics. One can compute the sensitivity curve for LISA or other future gravitational wave detectors (e.g. TianQin's curve, or that of a custom instrument) and use it to compute the signal-to-noise ratio of a collection of sources. Furthermore, \lw{} provides tools to visualise all of the results with easy-to-use plotting functions. Finally, \lw{} is fully tested to check for consistency in the derivations described below.

Specifically, we implement the post-Newtonian expressions by \citet{Peters+1963} and \citet{Peters+1964} for the evolution of binary orbits due to the emission of gravitational waves, equations for the strain amplitudes and signal-to-noise ratios of binaries from various papers \citep[e.g.][]{Flanagan+1998, Finn+2000, Cornish2003, Barack+2004, Moore+2015} and approximations for the LISA and TianQin sensitivity curves given in \citet{Robson+2019} and \citet{Huang+2020} respectively.

>>>>

The planned space-based gravitational wave detector LISA (Laser Interferometer Space Antenna) will present an entirely new view of gravitational waves by focusing on lower frequencies ($10^{-5} < f / \unit{Hz} < 10^{-1}$) than ground-based detectors. This will enable the study of many new source classes including mergers of supermassive black holes \citep[e.g.][]{Begelman+1980, Klein+2016, Bellovary2019}, extreme mass ratio inspirals \citep[e.g.][]{Berti2006, Barack2007, Babak2017, Moore2017}, and cosmological GW backgrounds \citep[e.g.][]{Bartolo2016, Caprini2016, Caldwell2019}. However, this frequency regime is also of interest for the detection of local stellar-mass binaries during their inspiral phase. LISA is expected to detect Galactic stellar-origin binaries containing combinations of white dwarfs, neutron stars, and black holes, ranging from the numerous double white dwarf population, to the rare but loud double black hole population.

@@

\lw{} is an open-source Python package designed to streamline the process of making predictions of LISA detection rates for stellar-origin binaries such that it is as fast, reliable and simple as possible. With \lw{} one can evolve the orbits of a binary or a collection of binaries and calculate their strain amplitudes for any range of frequency harmonics. One can compute the sensitivity curve for LISA or other future gravitational wave detectors (e.g.\ TianQin's curve, or that of a custom instrument) and use it to compute the signal-to-noise ratio of a collection of sources. Furthermore, \lw{} provides tools to visualise all of the results with easy-to-use plotting functions. Finally, \lw{} is fully tested to check for consistency in the derivations described below.

Specifically, we implement the post-Newtonian expressions by \citet{Peters+1963} and \citet{Peters+1964} for the evolution of binary orbits due to the emission of gravitational waves, equations for the strain amplitudes and signal-to-noise ratios of binaries from various papers \citep[e.g.][]{Flanagan+1998, Finn+2000, Cornish2003, Barack+2004, Moore+2015} and approximations for the LISA and TianQin sensitivity curves given in \citet{Robson+2019} and \citet{Huang+2020} respectively. \referee{The post-Newtonian expressions are approximately of order 0.5 as they account for the orbital evolution from energy loss due to the emission of GWs, but don’t include other effects such as spin-orbit coupling. We find this is an excellent approximation for stellar-origin sources in the LISA band as any higher order terms are either zero or negligible.}
